The Splashback Radius of Planck-SZ clusters

### 2018-11-01 Paper_new/RSP_paper_new.tex
Added one halo derivative profiles. Made small changes

\citet{umetsu2017lensing} used Xray selected clusters to look for
the splashback radius using the weak lensing signal, however
stacking issues and the low signal-to-noise ratio remains a significant
hurdle. \citet{chang2017splashback} found evidence for the splashback
radius in the weak lensing signal, but their analysis was again done
using optically selected clusters in the Dark energy survey
\citep{dark2005dark}. Regardless of the projection issues present in
the optical cluster catalog, there is some inherent circularity
present in the logic of using photometric galaxies to select
clusters as over-densities in a given aperture, and then using the
same photometric sample of galaxies to look for the splashback
radius. There is a possibility that the aperture used to select the
cluster catalogs could be imprinted in a non-trivial way on the measured number
density profiles of clusters. {\bf Add citation to the more recent Xray paper
by Henk Hoekstra and collaborators as well.}

@@

We use the Davis-Peebles estimator \citep{davis1983survey} to
compute the cross-correlation between our galaxy clusters and
galaxies. This estimator can be written as
\begin{equation}
\xi_{\rm 2D}(R) = \frac{\rm D_1D_2-R_1D_2}{\rm R_1D_2}\,
\end{equation}
where D$_1$D$_2$ and R$_1$D$_2$ are the normalized numbers of
cluster-galaxy pairs and cluster randoms-galaxy pairs at a given
comoving projected separation $R$. The subtraction of the signal
around random cluster positions gets rid of the uncorrelated pairs and allows
us to estimate the projected cross-correlation. The uncertainty in the galaxy
distribution masks prevents us from using the Landy-Szalay estimator. {\bf Cite
Landy-Szalay paper.}

@@

In order to compute the color cut to be used, we first match spectroscopic
galaxies in the Sloan digital Sky Survey ({\bf which data release?}) to their
Pan-STARRS photometry. We bin these galaxies in narrow redshift bins and
produce a histogram of the $g_{\mathrm{P1}}-r_{\mathrm{P1}}$ colors of the
galaxies in each bin. Due to the presence of the $4000\AA$-break present in the
red galaxy population, the two galaxy populations separate in such histograms
into two populations and we fit a double Gaussian distribution to it.  Based on
the fitted distribution we put a cut in color to exclude a contamination of the
blue population by incorrectly characterized red galaxies with a confidence of
3$\sigma$. We repeat this procedure for each redshift bin, to obtain a redshift
dependent color cut that separates the red from the blue population minimizing
the contamination of the blue population by red galaxies. Although being much
simpler and faster than calculating the individual k-corrections for each
galaxy, this procedure has its own shortcomings as discussed in
Appendix~\ref{sec:Errors}, in particular due to the photometric errors of red
galaxies. We would therefore exercise appropriate caution while interpreting
the results obtained by dividing galaxies in to color bins.

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.85]{3D_graphs_with_spline_no_mc.pdf}
\caption{The estimates of the three-dimensional cross-correlation
signals are shown in the leftmost column. The colored curves show the
model fits of the functional form in Equation~\ref{eq:model}. The magnitude
limit applied on the galaxy catalog is indicated in each row along with 
the signal-to-noise ratio (SNR). The vertical, shaded regions indicate the 
estimates of the splashback radii as estimated from the corresponding minima 
of the derivative profiles, which are shown in the column in the middle. The
rightmost column shows the second derivative profiles in matching color.
For comparison, the location of the $r_{\mathrm{200m}}$ radius as calculated 
from the average cluster sample properties is indicated by the black, dotted
lines.}
   \label{fig:3D_graphs} 
\end{figure*}

@@

\begin{figure*}
\centering{
  \includegraphics[scale=0.8]{derivatives_with_spline_no_mc.pdf}}
\caption{We show the distributions of the logarithmic derivatives of
the three-dimensional cross-correlation signals at the location of
the splashback radius. In the left panel the distributions of the
derivatives of the full profiles are shown, whereas only the inner
halo term (namely $\rho_{\mathrm{in}}f_{\mathrm{trans}}$) is
considered in the right panel.}
   \label{fig:derivatives} 
\end{figure*}
\begin{figure*}
\centering{
    \includegraphics[scale=0.8]{splashback_with_spline_no_mc.pdf}}
\caption{Shown are the posterior distributions of the inferred
locations of steepest slope of the two- and three dimensional
cross-correlation signals. On the left side the distributions of the
locations of the two-dimensional steepening feature
($R_{\mathrm{sp}}^{\mathrm{2D}}$) are shown, whereas the
distributions of the three-dimensional counterparts
($r_{\mathrm{sp}}^{\mathrm{3D}}$) are shown on the right. The
magnitude limit applied to the galaxy catalogs is indicated for each
distribution. Note that the colors of the distributions match with
the colors used in Figure~\ref{fig:2D_graphs} and
Figure~\ref{fig:3D_graphs}. For comparison, the location of the virial
radius $r_{\mathrm{200m}}$ as calculated from the average cluster
sample properties is indicated by the black, dotted lines.}
   \label{fig:splashback} 
\end{figure*}
\begin{figure}
    \includegraphics[scale=0.65]{color_separated_2D_with_spline_no_mc.pdf}
\caption{The two-dimensional cross-correlations
of the red and blue galaxy populations as inferred by
cross-correlating the cluster sample with the color-separated
subsamples that were extracted from the PS 21.5 galaxy catalog are
shown in the left panel, whereas the associated derivative profiles
are shown on the right. The vertical, shaded bands indicate the
locations of steepest slope of the cross-correlation signal as
inferred from the two subsamples, whereas the black, dashed lines
indicate the upper and lower bounds of the same feature but as
estimated from the full PS 21.5 galaxy catalog. }
   \label{fig:color_curve_2D} 
\end{figure}
\begin{figure*}
    \includegraphics[scale=0.75]{color_separated_3D_with_spline_no_mc.pdf}
\caption{We show the three-dimensional cross-correlations as
inferred by cross-correlating the cluster sample with the
color-separated galaxy subsamples, as well as the corresponding
splashback radii (indicated by the colored, vertical bands) and
their first and second order derivative profiles. Here, the black, dashed lines indicate
upper and lower bounds of the three-dimensional splashback radius
corresponding to the full PS 21.5 galaxy sample.}
   \label{fig:color_curve_3D} 
\end{figure*}
\begin{figure}
    \includegraphics[width= \columnwidth]{color_fraction_with_spline_no_mc.pdf}
\caption{Fractional contributions of the red and blue galaxy
populations to the total galaxy number density. We show the
locations of the splashback radii associated with the two subsamples
with the colored, vertical bands and the upper and lower bounds of
the splashback radius obtained from the full PS 21.5 catalog with
the black, dashed lines.}
   \label{fig:color_fraction} 
\end{figure}
The right panels in Figure~\ref{fig:2D_graphs} show the
corresponding analytical derivatives of the two-dimensional
cross-correlations and the 68\% confidence interval based on our
model fits. The logarithmic derivatives show a distinct steepening
feature at around $1.3~\mpch$. The figure shows that the
location of the steepest slope does not change appreciably when the
magnitude limit is changed, even though we use a sample which is one
magnitude brighter than $M_{\rm i}-5\log h=-19.44$. The 68\%
confidence interval of the location of the steepest slope is
indicated by the vertical, shaded region. We also show the location
of the virial radius $r_{\mathrm{200m}}$ based on the Planck SZ mass estimate 
as a black, dotted line in each panel.

@@

Following \citet{baxter2017halo}, we also present the values of the
logarithmic derivatives at the location of the steepest slope for
the total three-dimensional cross-correlations, as well as those for
the inner halo term in the left and right hand panels of
Figure~\ref{fig:derivatives}, respectively. The logarithmic slope of the 
cross-correlation is significantly steeper than $-3$ at the location 
of the splashback radius, making it difficult to
be reproduced by classical fitting functions like the NFW profile,
which reach such slopes only asymptotically and even that only
without the presence of the outer 2-halo term. This provides
evidence for the existence of the splashback feature.

>>>>

\citet{umetsu2017lensing} as well as \citet{contigiani2018weak} 
used Xray selected clusters to look for the splashback radius 
using the weak lensing signal, however
stacking issues and the low signal-to-noise ratio remains a significant
hurdle for both of them. \citet{chang2017splashback} found evidence for the splashback
radius in the weak lensing signal, but their analysis was again done
using optically selected clusters in the Dark energy survey
\citep{dark2005dark}. Regardless of the projection issues present in
the optical cluster catalog, there is some inherent circularity
present in the logic of using photometric galaxies to select
clusters as over-densities in a given aperture, and then using the
same photometric sample of galaxies to look for the splashback
radius. There is a possibility that the aperture used to select the
cluster catalogs could be imprinted in a non-trivial way on the measured number
density profiles of clusters. {\bf Add citation to the more recent Xray paper
by Henk Hoekstra and collaborators as well.}

@@

We use the Davis-Peebles estimator \citep{davis1983survey} to
compute the cross-correlation between our galaxy clusters and
galaxies. This estimator can be written as
\begin{equation}
\xi_{\rm 2D}(R) = \frac{\rm D_1D_2-R_1D_2}{\rm R_1D_2}\,
\end{equation}
where D$_1$D$_2$ and R$_1$D$_2$ are the normalized numbers of
cluster-galaxy pairs and cluster randoms-galaxy pairs at a given
comoving projected separation $R$. The subtraction of the signal
around random cluster positions gets rid of the uncorrelated pairs and allows
us to estimate the projected cross-correlation. The uncertainty in the galaxy
distribution masks prevents us from using the Landy-Szalay estimator \citep{landy1993bias}. 
{\bf Cite Landy-Szalay paper.}

@@

In order to compute the color cut to be used, we first match spectroscopic
galaxies in the Sloan digital Sky Survey (DR 13) ({\bf which data release?}) to their
Pan-STARRS photometry. We bin these galaxies in narrow redshift bins and
produce a histogram of the $g_{\mathrm{P1}}-r_{\mathrm{P1}}$ colors of the
galaxies in each bin. Due to the presence of the $4000\AA$-break present in the
red galaxy population, the two galaxy populations separate in such histograms
into two populations and we fit a double Gaussian distribution to it.  Based on
the fitted distribution we put a cut in color to exclude a contamination of the
blue population by incorrectly characterized red galaxies with a confidence of
3$\sigma$. We repeat this procedure for each redshift bin, to obtain a redshift
dependent color cut that separates the red from the blue population minimizing
the contamination of the blue population by red galaxies. Although being much
simpler and faster than calculating the individual k-corrections for each
galaxy, this procedure has its own shortcomings as discussed in
Appendix~\ref{sec:Errors}, in particular due to the photometric errors of red
galaxies. We would therefore exercise appropriate caution while interpreting
the results obtained by dividing galaxies in to color bins.

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.85]{3D_graphs_with_spline_no_mc.pdf}
\caption{The estimates of the three-dimensional cross-correlation
signals are shown in the leftmost column. The colored curves show the
model fits of the functional form in Equation~\ref{eq:model}. The magnitude
limit applied on the galaxy catalog is indicated in each row along with 
the signal-to-noise ratio (SNR). The vertical, shaded regions indicate the 
estimates of the splashback radii as estimated from the corresponding minima 
of the derivative profiles, which are shown in the column in the middle. The
rightmost column shows the second derivative profiles in matching color.
The grey curves show the corresponding, estimated first and second derivatives of the 
one-halo term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14}).
For comparison, the location of the $r_{\mathrm{200m}}$ radius as calculated 
from the average cluster sample properties is indicated by the black, dotted
lines.}
   \label{fig:3D_graphs} 
\end{figure*}

@@

\begin{figure*}
\centering{
  \includegraphics[scale=0.8]{derivatives_with_spline_no_mc.pdf}}
\caption{We show the distributions of the logarithmic derivatives of
the three-dimensional cross-correlation signals at the location of
the splashback radius. In the left panel the distributions of the
derivatives of the full profiles are shown, whereas only the inner
halo term (namely $\rho_{\mathrm{in}}f_{\mathrm{trans}}$) is
considered in the right panel.}
   \label{fig:derivatives} 
\end{figure*}
\begin{figure*}
\centering{
    \includegraphics[scale=0.8]{splashback_with_spline_no_mc.pdf}}
\caption{Shown are the posterior distributions of the inferred
locations of steepest slope of the two- and three dimensional
cross-correlation signals. On the left side the distributions of the
locations of the two-dimensional steepening feature
($R_{\mathrm{sp}}^{\mathrm{2D}}$) are shown, whereas the
distributions of the three-dimensional counterparts
($r_{\mathrm{sp}}^{\mathrm{3D}}$) are shown on the right. The
magnitude limit applied to the galaxy catalogs is indicated for each
distribution. Note that the colors of the distributions match with
the colors used in Figure~\ref{fig:2D_graphs} and
Figure~\ref{fig:3D_graphs}. For comparison, the location of the virial
radius $r_{\mathrm{200m}}$ as calculated from the average cluster
sample properties is indicated by the black, dotted lines.}
   \label{fig:splashback} 
\end{figure*}
\begin{figure}
    \includegraphics[scale=0.65]{color_separated_2D_with_spline_no_mc.pdf}
\caption{The two-dimensional cross-correlations
of the red and blue galaxy populations as inferred by
cross-correlating the cluster sample with the color-separated
subsamples that were extracted from the PS 21.5 galaxy catalog are
shown in the left panel, whereas the associated derivative profiles
are shown on the right. The vertical, shaded bands indicate the
locations of steepest slope of the cross-correlation signal as
inferred from the two subsamples, whereas the black, dashed lines
indicate the upper and lower bounds of the same feature but as
estimated from the full PS 21.5 galaxy catalog. }
   \label{fig:color_curve_2D} 
\end{figure}
\begin{figure*}
    \includegraphics[scale=0.76]{color_separated_3D_with_spline_no_mc.pdf}
\caption{We show the three-dimensional cross-correlations as
inferred by cross-correlating the cluster sample with the
color-separated galaxy subsamples, as well as the corresponding
splashback radii (indicated by the colored, vertical bands) and
their first and second order derivative profiles. The light-colored 
curves show the corresponding first and second derivatives of the one-halo
term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14})
Here, the black, dashed lines indicate upper and lower bounds of the 
three-dimensional splashback radius corresponding to the full PS 21.5 galaxy sample.}
   \label{fig:color_curve_3D} 
\end{figure*}
\begin{figure}
    \includegraphics[width= \columnwidth]{color_fraction_with_spline_no_mc.pdf}
\caption{Fractional contributions of the red and blue galaxy
populations to the total galaxy number density. We show the
locations of the splashback radii associated with the two subsamples
with the colored, vertical bands and the upper and lower bounds of
the splashback radius obtained from the full PS 21.5 catalog with
the black, dashed lines.}
   \label{fig:color_fraction} 
\end{figure}
The right panels in Figure~\ref{fig:2D_graphs} show the
corresponding analytical derivatives of the two-dimensional
cross-correlations and the 68\% confidence interval based on our
model fits. The logarithmic derivatives show a distinct steepening
feature at around $1.3~\mpch$. The figure shows that the
location of the steepest slope does not change appreciably when the
magnitude limit is changed, even though we use a sample which is one
magnitude brighter than $M_{\rm i}-5\log h=-19.44$. The 68\%
confidence interval of the location of the steepest slope is
indicated by the vertical, shaded region. We also show the location
of the virial radius $r_{\mathrm{200m}}$ based on the Planck SZ mass estimate 
as a black, dotted line in each panel.

@@

Following \citet{baxter2017halo}, we also present the values of the
logarithmic derivatives at the location of the steepest slope for
the total three-dimensional cross-correlations, as well as those for
the inner halo term in the left and right hand panels of
Figure~\ref{fig:derivatives}, respectively. The logarithmic slope of the 
cross-correlation is significantly steeper than $-3$ at the location 
of the splashback radius, making it difficult to
be reproduced by classical fitting functions like the NFW profile,
which reach such slopes only asymptotically and even that only
without the presence of the outer 2-halo term. This provides
evidence for the existence of the splashback feature. From the second
derivative profiles we also note that the drop in the cross-correlation
signal is very sharp and happens within a factor 2 in radius.
